Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-14 related.tex
boom, paragraph

{\color{red}[[Ryan to write]]

A paragraph on redundancy and its connection with predictability  how it is approximated and how these estimates are bad

e.g., Redundancy estimates rely heavily on either estimating the generating partition which is hard to do *and* estimating the full positive Lyapunov spectrum which is hard to do for noisy systems and impossible for systems that are not deterministic.

Also please edit / look at the paragrpah proceeding and following this one
}

>>>>

It seems desirable to quantify how much information prior observations of a system lend to future values, the \emph{mututal information} between, say, the past $n$ observations at the current one. This quantity is known as the \emph{redundancy}. For example, this would be the amount of information that delay coordinate embedding captures. For IID random processes,  the redundancy is zero, since all observations are independent of each other. On the other hand, any deterministic system --- including chaotic ones --- can be perfectly predicted if observed for long enough~\cite{{weigend-book}. Unfortunately it is quite difficult to estimate the redundancy in practice for an arbitrary, real-valued time series. Doing so requires knowing either the Kolmolgorov-Sinai entropy or the Lyapunov spectrum. As discussed above, computing the entropy is quite difficult. Similarly, computing the Lyapunov exponent from an arbitrary time series is also fraught with difficulties, and is effectivly impossible if there is significant noise or the generating system is stochastic.
